---
title: "Homework 6"
author: "George Tapia (A20450857)"
date: "2023-04-03"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Part 2.1(b)
```{r}
library(arules)

working_directory <- getwd()
set_working_directory <- working_directory
```

### Transactions Equal to 1k (minsup = 0.01, minconf = 0.85)
```{r}
tr1k <- read.transactions("tr-1k-canonical.csv", sep = ",")
frequent_item_sets_tr1k <- apriori(tr1k, parameter = list(support = 0.01, target = "frequent itemsets"))
rules_for_tr1k <- apriori(tr1k, parameter = list(support = 0.01, conf = 0.85, target = "rules"))
```
#### Frequent Itemsets (Tr1k)
```{r}
print(inspect(sort(frequent_item_sets_tr1k, decreasing = T, by="count")))
```
#### Rules (Tr1k)
```{r}
print(inspect(sort(rules_for_tr1k, by="confidence")))
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
### Transactions Equal to 5k (minsup = 0.01, minconf = 0.85)
```{r}
tr5k <- read.transactions("tr-5k-canonical.csv", sep = ",")
frequent_item_sets_tr5k <- apriori(tr5k, parameter = list(support = 0.01, target = "frequent itemsets"))
rules_for_tr5k <- apriori(tr5k, parameter = list(support = 0.01,conf = 0.85, target = "rules"))
```
#### Frequent Itemsets (Tr5k)
```{r}
print(inspect(sort(frequent_item_sets_tr5k, decreasing = T, by="count")))
```

#### Rules (Tr5k)
```{r}
print(inspect(sort(rules_for_tr5k, by="confidence")))
```

### Transactions Equal to 20k (minsup = 0.01, minconf = 0.85)
```{r}
tr20k <- read.transactions("tr-20k-canonical.csv", sep = ",")
frequent_item_sets_tr20k <- apriori(tr20k, parameter = list(support = 0.01, target = "frequent itemsets"))
rules_for_tr20k <- apriori(tr20k, parameter = list(support = 0.01, conf = 0.85, target = "rules"))
```
#### Frequent Itemsets (Tr20k)
```{r}
print(inspect(sort(frequent_item_sets_tr20k, decreasing = T, by="count")))
```

#### Rules (Tr20k)
```{r}
print(inspect(sort(rules_for_tr20k, by="confidence")))
```

### Transactions Equal to 75k (minsup = 0.01, minconf = 0.85)
```{r}
tr75k <- read.transactions("tr-75k-canonical.csv", sep = ",")
frequent_item_sets_tr75k <- apriori(tr75k, parameter = list(support = 0.01, target = "frequent itemsets"))
rules_for_tr75k <- apriori(tr75k, parameter = list(support = 0.01, conf = 0.85, target = "rules"))
```

#### Frequent Itemsets (Tr75k)
```{r}
print(inspect(sort(frequent_item_sets_tr75k, decreasing = T, by="count")))
```

#### Rules (Tr75k)
```{r}
print(inspect(sort(rules_for_tr75k, by="confidence")))
```

### Part 2.1(c)
#### Given the comparison of the rules, it seems like the amount of rules was slightly lower for the 1k dataset.
#### However, for the other datasets, the number of rules did not differ. Additionally, the rules are suprisingly the same.

### Part 2.1(d)
#### (i)  {Coffee Eclair}
#### (ii) {Apple Croissant, Apple Danish, Apple Tart, Cherry Soda}