---
title: "Homework 9"
author: "George Tapia (A20450857)"
date: "2023-05-01"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
---

### Part 2.1 (a)

#### (i) 
#### We should remove the name before we do clustering since this value is unique,
#### it will not provide us with any useful insights
#### After looking at the dataset, I noticed that, top canines (C) and bottom canines (c)
#### had similar values. I double checked by using a correlation plot.
#### since the two attributes are highly correlated. If we use both of them, this can lead us
#### to less reliable statistical inferences so it is best to remove one.
#### I will remove the bottom canine attribute (random choice- both are similar)

#### (ii)
#### Given that all the features don't have a significant difference, in terms of scale,
#### we do not need to standardize the data

#### (iii) (below, done through code)
```{r}
library(factoextra)
library(corrplot)
working_directory <- getwd()
setwd(working_directory)
mammals_old <- read.table("file19.txt", header = TRUE, skip = 20)


# the tooth pattern
#
#(I) the number of top incisors,
#(i) bottom incisors, 
#(C) top canines, 
#(c) bottom canines, 
#(P) top premolars, 
#(p) bottom premolars, 
#(M) top molars
#(m) bottom molars.

correlation_matrix <- cor(mammals_old[2:9])
corrplot(correlation_matrix, method = "color", type = "upper", order = "hclust", addCoef.col = "black", tl.col = "black", tl.srt = 45)


mammals <- mammals_old[,-c(5)]
print(mammals)

```

### Part 2.1 (b) (REMOVED ATTRIBUTE - BOTTOM CANINES (c))
#### (i) Using the silhoutte method, we were able to determine that the 
####  optimal number of clusters is 8. This means that each
####  number of cluster's were evaluated to determine the number
####  that gave the best results, with respect to how well they were clustered.
```{r}
fviz_nbclust(mammals[,-1], kmeans, method = "silhouette")
```



#### (ii)
```{r}
km.mammals <- kmeans(mammals[,-1], 8, nstart = 10)
fviz_cluster(km.mammals, mammals[,-1])
```

#### (iii) 1, 9, 12, 7, 8, 11, 10, 8  (respectively)
```{r}
cluster_one <- mammals[which(km.mammals$cluster == 1),]
cluster_two <- mammals[which(km.mammals$cluster == 2),]
cluster_three <- mammals[which(km.mammals$cluster == 3),]
cluster_four <- mammals[which(km.mammals$cluster == 4),]
cluster_five <- mammals[which(km.mammals$cluster == 5),]
cluster_six <- mammals[which(km.mammals$cluster == 6),]
cluster_seven <- mammals[which(km.mammals$cluster == 7),]
cluster_eight <- mammals[which(km.mammals$cluster == 8),]

cat(" Number of observations in Cluster 1: ", nrow(cluster_one), "\n",
    "Number of observations in Cluster 2: ", nrow(cluster_two), "\n",
    "Number of observations in Cluster 3: ", nrow(cluster_three), "\n",
    "Number of observations in Cluster 4: ", nrow(cluster_four), "\n",
    "Number of observations in Cluster 5: ", nrow(cluster_five), "\n",
    "Number of observations in Cluster 6: ", nrow(cluster_six), "\n",
    "Number of observations in Cluster 7: ", nrow(cluster_seven), "\n",
    "Number of observations in Cluster 8: ", nrow(cluster_eight), "\n")


```

#### (iv) 551.94
```{r}
km.mammals <- kmeans(mammals[,-1], 8, nstart = 10)
print("Total SSE:")
print(km.mammals$totss)
```

#### (v) 0 15.55  8.83 17.71  6.38  5.45  5.7  3.88 (respectively)
```{r}

print(km.mammals)

```

#### (vi)
#### Cluster 1: Only one animal so can't really deduce anything from this cluster yet.
#### However, armadillos have no canines, after research, so it can be meaningful.
```{r}
print(cluster_one[, 1])
```
#### Cluster 2: All of these animals are ugly and creepy, except for the bear.
#### All share a similar characteristic. They all have a lot of fur.
```{r}
print(cluster_two[, 1])
```
#### Cluster 3: Several of these animals are fast, and use their senses to hunt their prey
```{r}
print(cluster_three[, 1])
```
#### Cluster 4: martens, fishers, and wolverines are predators that feed on smaller mammals and bird,
#### while fur seals, sea lions, walruses, and elephant seals primarily consume fish 
#### and other marine animals. This cluster is meh.
```{r}
print(cluster_four[, 1])
```
#### Cluster 5: most of these animals are bats so this is a great grouping
```{r}
print(cluster_five[, 1])
```
#### Cluster 6: The majority of these animals are rodents so this is a good cluster as well
```{r}
print(cluster_six[, 1])
```
#### Cluster 7: All of the animals, except one, are Ungulates.
#### Ungulates are characterized by their hooved feet, which typically have an even or odd number of toes.
#### Coyotes are not ungulates, but they share a common habitat with many ungulates.
#### Great grouping!
```{r}
print(cluster_seven[, 1])
```
#### Cluster 8: Most of these are rodents so this is a good grouping as well
```{r}
print(cluster_eight[, 1])
```

#### In summary, Cluster 1 consists of a single animal, the armadillo, which is notable for its lack of 
#### canines, making it an interesting case. Cluster 2 is made up of animals with abundant fur, although they
#### vary in species and are subjectively seen as "ugly and creepy" with the exception of the bear. Cluster 3
#### contains several swift animals that rely on their senses while hunting. Cluster 4 encompasses both land
#### predators and marine mammals. Cluster 5 predominantly includes bats,
#### demonstrating a cohesive grouping. Clusters 6 and 8 mostly consist of rodents, indicating appropriate
#### groupings. Cluster 7 is largely made up of ungulates, with the exception of the coyote, which inhabits
#### similar environments as many ungulates. Overall, great findings!


### Part 2.2 
#### (a) I looked at the dataset. Since I can't tell what the attributes are, it would be best
#### to scale the data to prevent any distortion.
```{r}
library(dbscan)
library(fpc)
s1 <- read.csv("s1.csv", header = TRUE)
s1.scaled <- scale(s1)


```

#### (b)(i - ii)
#### I see 15 notable clusters from the chart. Though, all of them are not well seperated (only one).
```{r}
plot(s1, main="Raw points")
```



#### (c)(i - iii)
#### I would only used 15 clusters since the plot "Raw points" shows 15 distinguishable clusters.
#### The plot shows me that 15 clusters will be a good classification pattern to analyze for insights.
```{r}
fviz_nbclust(s1.scaled, kmeans, method = "wss", k.max = 20)  + geom_vline(xintercept = 4, linetype = 2)
```

```{r}
fviz_nbclust(s1.scaled, kmeans, method = "silhouette", k.max = 20)
```

#### (d)(i)
#### (ii) I think the clustering is great since we have 15 clusters.
#### All the data looks clustered in a good way that will give us valuable insights.
```{r}
km.s1.scaled <- kmeans(s1.scaled, 15, nstart = 10)
fviz_cluster(km.s1.scaled, s1)
```

#### (e) 
```{r}

kNNdistplot(s1.scaled, k = 8) 

dbscan.s1.scaled <- fpc::dbscan(s1.scaled, eps = 0.095, MinPts = 8)
fviz_cluster(dbscan.s1.scaled, s1, stand = FALSE, ellipse = F, geom = "point")

```

#### (i)
#### a value of 8 would be best becuase the clusters look more dense.
#### (ii) 
#### At minPts = 8, eps = 0.095, there are 15 clusters.