---
title: "CS 422 Homework 4"
author: "George Tapia (A20450857)"
date: "2023-03-04"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##### Section 2.1

### Part A
```{r}
setwd("C:/Users/gtapi/OneDrive/Documents/R_Files")
hotel_bookings <- read.csv("hotel_bookings.csv")

table(hotel_bookings$hotel)
```
### Part B
```{r}
class_frequency <- table(hotel_bookings$is_canceled)

# retrieving number of canceled customers
canceled_count <- class_frequency[["1"]]

# retrieving number of non canceled customers
not_canceled_count <- class_frequency[["0"]]

cat("Number of guests who canceled reservation:", canceled_count, "\n",
    "Number of guests who did not cancel the reservation:", not_canceled_count)
```
### Part C
```{r}
customer_type_frequency <- table(hotel_bookings$customer_type)


# returning the index of the max number of customers for the most number of reservations
max_index <- which.max(customer_type_frequency)

# grabbing the customer type
most_common_customer_type <- names(customer_type_frequency)[max_index]

# grabbing the number of reservations from the frequency table
number_of_reservations <- customer_type_frequency[max_index]

output_string <- paste("Customer type with the most reservations is", most_common_customer_type, ", with", 
                       number_of_reservations, "reservations")
print(output_string)
```
### Part D
```{r}

parking_space_frequency <- table(hotel_bookings$required_car_parking_spaces)

# take the attributes(required number of parking spaces) and turn it into a vector
required_parking_spaces <- as.numeric(names(parking_space_frequency))

# iterate through the vector, and retrieve the max number of required parking spaces
max_number_of_required_spaces <- max(required_parking_spaces)

# iterate through the vector and grab the index of the minimum required of parking spaces
max_required_parking_spaces_index <- which.max(required_parking_spaces)

# grab the number of customers who requested the max number of parking spaces from the frequency table
number_of_customers_with_max_spaces <-parking_space_frequency[max_required_parking_spaces_index]


output_string <- paste(number_of_customers_with_max_spaces, " customers required the most number of parking spaces (", max_number_of_required_spaces, ").", sep = "")

print(output_string)

matching_rows <- hotel_bookings[hotel_bookings$required_car_parking_spaces == max_number_of_required_spaces, ]
print(matching_rows)
```
### Part E
```{r}
parking_space_frequency <- table(hotel_bookings$required_car_parking_spaces)

# take the attributes (required number of parking spaces) and turn it into a vector
required_parking_spaces <- as.numeric(names(parking_space_frequency))

# iterate through the vector, and retrieve the minimum number of required parking spaces
min_number_of_required_spaces <- min(required_parking_spaces)

# iterate through the vector and grab the index of the minimum required of parking spaces
min_required_parking_spaces_index <- which.min(required_parking_spaces)

# grab the number of customers who requested the min number of parking spaces from the frequency table
number_of_customers_with_min_spaces <-parking_space_frequency[min_required_parking_spaces_index]


output_string <- paste(number_of_customers_with_min_spaces, " customers required the least number of parking spaces (", min_number_of_required_spaces, ").", sep = "")

print(output_string)

#matching_rows <- hotel_bookings[hotel_bookings$required_car_parking_spaces == min_number_of_required_spaces, ]
#print(matching_rows)

```
### Part F
```{r}
# count the number of people who expressed a preference for a room type during reservation
# using the is.na function to count the number of cells that contain a value

#num_null <- sum(is.na(hotel_bookings$reserved_room_type))
num_preferred <- sum(!is.na(hotel_bookings$reserved_room_type))

# count the number of people who were assigned their preferred room type during check-in
# the equality function returns a vector of true and false values
# the sum function returns the sum of the true values only
num_assigned_pref <- sum(hotel_bookings$assigned_room_type == hotel_bookings$reserved_room_type)

# calculate the percentage of people who were assigned their preferred room type
percentage <- 100 * num_assigned_pref / num_preferred

cat(sprintf("%.2f%% of the people who expressed a room preference during reservation got the room during check-in.", percentage))

```
```{r}
# create separate data frames for city and resort hotels
# create a subset where hotel attribute equals type of hotel
city_hotels <- subset(hotel_bookings, hotel == "City Hotel")
resort_hotels <- subset(hotel_bookings, hotel == "Resort Hotel")

top_10_city_hotel_countries <- head(sort(table(city_hotels$country), decreasing = TRUE), 10)

# noticed a NULL value here (anamoly)
top_11_resort_hotel_countries <- head(sort(table(resort_hotels$country), decreasing = TRUE), 11)

# removing the NULL value from the 11 resort hotel countries to make it 10
if("NULL" %in% names(top_11_resort_hotel_countries)){
  top_10_resort_hotel_countries <- top_11_resort_hotel_countries[names(top_11_resort_hotel_countries) != "NULL"]
}

barplot(sort(top_10_city_hotel_countries, increasing = TRUE), 
        col = rainbow(10),
        ylim = c(0, 9000),
        main = "Top 10 Countries of Origin for City Hotels")

barplot(sort(top_10_resort_hotel_countries, increasing = TRUE), 
        col = rainbow(10),
        ylim = c(0, 7000),
        main = "Top 10 Countries of Origin for Resort Hotels")


most_visitors <- names(which.max(table(hotel_bookings$country)))
cat("The country with the most visitors is", most_visitors, "\n")
```
###### Section 2.2 (Part A)

### Cleaning the data (combining arrival day, month, and year into one column. Then generating a table to see where all the null values are at)
```{r}
set.seed(1122)


# Load the required packages
library(xgboost)

# the first thing that we need to do is clean the data
# we can minimize the number of columns by combining the arrival month, day, and year into a single column
# I will first change the values of the arrival months to a numeric format so I can have the concatenation match the reservation date status
month_names <- c("January" = "01", "February" = "02", "March"="03", "April"="04", "May"="05", "June"="06",
                 "July"="07", "August"="08", "September"="09", "October"="10", "November"="11", "December"="12")

# just adding the leading zeros to generate a date format like the reservation status date
hotel_bookings$arrival_date_month <- month_names[hotel_bookings$arrival_date_month]
hotel_bookings$arrival_date_day_of_month <- sprintf("%02d", hotel_bookings$arrival_date_day_of_month)

# combining the values to form the single column
arrival_date <- paste(hotel_bookings$arrival_date_year, hotel_bookings$arrival_date_month, hotel_bookings$arrival_date_day_of_month, sep ="-")

hotel_bookings$arrival_date <- arrival_date
hotel_bookings$arrival_date_year <- NULL
hotel_bookings$arrival_date_month <- NULL
hotel_bookings$arrival_date_day_of_month <- NULL

# taking the strings and creating date objects out of them so we can perform comparison operations
hotel_bookings$reservation_status_date <- as.Date(hotel_bookings$reservation_status_date)
hotel_bookings$arrival_date <- as.Date(hotel_bookings$arrival_date)


check_values <- c("Undefined", "undefined", "NULL", "", NA)

# finding the sum of missing values for each column
missing_values <- sapply(hotel_bookings, function (x) sum(x %in% check_values))
missing_percentages <- sapply(hotel_bookings, function (x) mean(x %in% check_values) * 100)

# displaying the tables of missing values
knitr::kable(missing_values, align = "c")
knitr::kable(missing_percentages, align = "c")

```

### Children column adjusted
```{r}
# taking care of the missing children columns first
missing_children_indices <- which(is.na(hotel_bookings$children))

# taking the mean of the children so that I can fill in the missing values with the mean value
# taking this approach since it is a very small amount of missing values
mean_of_children <- as.integer(floor(mean(hotel_bookings$children, na.rm = TRUE)))

# filling in the missing values with the mean value
hotel_bookings[missing_children_indices,]$children <- mean_of_children
```


### Meal missing values adjusted
```{r}
# reference https://stackoverflow.com/questions/2547402/how-to-find-the-statistical-mode
find_mode <- function(x) {
  u <- unique(x)
  tab <- tabulate(match(x, u))
  u[tab == max(tab)]
}
#---------------------------------------------------------------------------------------

missing_meal_indices <- which(hotel_bookings$meal == "Undefined")
mode_of_meals <- find_mode(hotel_bookings$meal)
hotel_bookings[missing_meal_indices,]$meal <- mode_of_meals
```

### Country missing values adjusted
```{r}
missing_country_indices <- which(hotel_bookings$country == "NULL")
mode_of_country <- find_mode(hotel_bookings$country)
hotel_bookings[missing_country_indices,]$country <- mode_of_country
```

### Market segment missing values adjusted
```{r}
missing_market_indices <- which(hotel_bookings$market_segment == "Undefined")
mode_of_market <-find_mode(hotel_bookings$market_segment)
hotel_bookings[missing_market_indices,]$market_segment <- mode_of_market
```

### Distribution channel missing values adjusted
```{r}
missing_distribution_indices <- which(hotel_bookings$distribution_channel == "Undefined")
mode_of_distribution <- find_mode(hotel_bookings$distribution_channel)
hotel_bookings[missing_distribution_indices,]$distribution_channel <- mode_of_distribution
```

### Agent missing values adjustment by mode of each country
```{r}
# grabbing the rows that have null for agent to see if there is a pattern
null_agent_indices<- which(hotel_bookings$agent == "NULL")

# changing all the string values in the agent column to integer values
hotel_bookings$agent <- as.integer(hotel_bookings$agent)

list_of_agents <- sort(unique(hotel_bookings[-null_agent_indices,]$agent), decreasing = TRUE)
agent_by_country <- aggregate(agent ~ country, data = hotel_bookings[-null_agent_indices,] , find_mode)

# extract the specific agent value for each country
agent_by_country$agent <- sapply(agent_by_country$agent, function(x) {
  if (length(x) == 1) {
    return(x[1])
  } else {
    return(sample(x, 1))
  }
})

# get indices of null agent values in hotel_bookings
null_agent_indices <- which(is.na(hotel_bookings$agent))

# iterate over rows with null agent values and fill in corresponding agent from agent_by_country
for (i in null_agent_indices) {
  matching_row <- which(agent_by_country$country == hotel_bookings[i, "country"])
  if (length(matching_row) > 0) {
    hotel_bookings[i, "agent"] <- as.integer(agent_by_country[matching_row, "agent"])
  }
  else{
    hotel_bookings[i, "agent"] <- as.integer(9)
  }
}

```
### Company column and arrival_date_week_number removed
```{r}
# since the company column has more than 70% of missing values, I will be removing it from the data set
# I will also remove the arrival date week number column since it is of no use, we have the arrival date already
hotel_bookings$company <- NULL
hotel_bookings$arrival_date_week_number <- NULL
```


### Data is finally clean, and ready to be processed
```{r}
check_values <- c("Undefined", "undefined", "NULL", "", NA)

# finding the sum of missing values for each column
missing_values <- sapply(hotel_bookings, function (x) sum(x %in% check_values))
missing_percentages <- sapply(hotel_bookings, function (x) mean(x %in% check_values) * 100)

# displaying the tables of missing values
knitr::kable(missing_values, align = "c")
knitr::kable(missing_percentages, align = "c")
```
### Splitting the data into test and training sets (set.seed was used in data cleaning phase)
```{r}

# Split the data into training and test sets
n_obs <- nrow(hotel_bookings)
train_idx <- sample(n_obs, 0.9 * n_obs)
train <- hotel_bookings[train_idx, ]
test <- hotel_bookings[-train_idx, ]

```

```{r}
library(rpart)
library(rpart.plot)
model <- rpart(is_canceled ~ deposit_type + previous_cancellations + arrival_date, method = "class", data = train)

rpart.plot(model, extra = 104, fallen.leaves = T, type = 4,
           main = "Rpart on Hotel data")
# now we can run our model on the test data set
pred <- predict(model, test, type = "class")
# create a table of actual vs predicted classes
conf_matrix <- table(test$is_canceled, pred)

# print the confusion matrix
print(conf_matrix)

# calculate accuracy
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
cat("Accuracy:", round(accuracy, 4), "\n")

# calculate error rate
error_rate <- 1 - accuracy
cat("Error rate:", round(error_rate, 4), "\n")

# calculate sensitivity
sensitivity <- conf_matrix[2, 2] / sum(conf_matrix[2, ])
cat("Sensitivity:", round(sensitivity, 4), "\n")

# calculate specificity
specificity <- conf_matrix[1, 1] / sum(conf_matrix[1, ])
cat("Specificity:", round(specificity, 4), "\n")

```